---
title: "PSF : R Package for Pattern Sequence based Forecasting Algorithm"
author: "Neeraj Dhanraj Bokde (neerajdhanraj@gmail.com)"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Introduction

The Algorithm Pattern Sequence based Forecasting (PSF) was first proposed by `Martinez Alvarez, et al., 2008` and then modified and suggested improvement by `Martinez Alvarez, et al., 2011`.  The technical detailes are mentioned in referenced articles. PSF algorithm consists of various statistical operations like:

- Data Normalization/ Denormalization
- Calculation of optimum Window size (W)
- Calculation of optimum cluster size (k)
- Pattern Sequence based Forecasting
- RMSE/MAE Calculation, etc..


## Examples

This section demonstrates the Functions used in PSF along with the examples.
The data used is from `iris` data set, provided by R.

```{r, echo=FALSE, results='asis'}
data(iris)
```


### Install library

Download the Package and install with instruction:

```{r}
library(PSF)
```



### `pred_for_w (data_in, w, k, next_val)`

This Function predict the values for given data, Window size (`W`) and Cluster size (`k`).

- `data_in` as Input data, in any format (data matrix data frame or vector). All variables should be numeric and NA values will get removed while execution.
- `w` as window size (Can be obtained with function `optimum_w`)
- `k` as cluster size for Kmeans (Can be obtained with function `optimum_k`)
- `next_val` as Integer number. It states the number of predicted values to be obtained.

```{r}
# Considering `data_in` = iris[1], `w` = 3, `k` = 4 and `next_val` = 5
pred_for_w(iris[1], 3, 4, 5)
```

This function returns the predicted values and corresponding graph plot. 
The graph region with Red Color is the original data, whereas the region with Blue color is corresponding to the predicted values.



### `pred_for_w_plot (data_in, w, k, next_val)`

This Function is similar to `pred_for_w()`, except this function is able to plot the predicted values.

```{r}
# Considering `data_in` = iris[1], `w` = 3, `k` = 4 and `next_val` = 5
pred_for_w_plot(iris[1], 3, 4, 5)
```



### `optimum_w (data_in, next_val)`

This Function calculates the optimum value of Window size `w` to be used in prediction function such that the `RMSE` and `MAE` values should be minimum.

- `data_in` as Input data, in any format (data matrix data frame or vector). All variables should be numeric and NA values will get removed while execution.
- `next_val` as Integer number. Such that last `next_val` numbers will be used for test of prediction accuracy.

```{r}
# Considering `data_in` = iris[1] and `next_val` = 5
optimum_w(iris[1], 5)
```

This Function returns the optimum value for `w`, it's corresponding `RMSE` values, prediction values and plots. The predicted values in `optimum_w()` function are different than those obtained in `pred_for_w()`, since, `optimum_w()` function removes last `next_val` numbers of integers and predicts that much values and compare them with original data.




### `optimum_k (data_in)`

This Function is to determine the Optimum value of Cluster size (`K`) based on Average silhouette method.

- `data_in` as Input data, in any format (data matrix data frame or vector). All variables should be numeric and NA values will get removed while execution.

```{r}
# Considering `data_in` = iris[1] 
optimum_k(iris[1])
```
This Function returns K value, an integer as optimum number clusters.




### `AUTO_PSF (data_in, next_val)`

This Function takes input data and autogenerate optimum Window size (`W`) and cluster size (`K`) and predicts the `next_val` numbers of future values

- `data_in` as Input data, in any format (data matrix data frame or vector). All variables should be numeric and NA values will get removed while execution.
- `next_val` as Integer number. It states the number of predicted values to be obtained.


```{r}
# Considering `data_in` = iris[1] and next_val = 3
AUTO_PSF(iris[1],3)
```

This function calculate suitable values of `w` and `k` such that mean erroe between original and predicted data should be minimum. And ultimately, forecasts the series with PSF algorithm methodology.




## References

Martínez-Álvarez, F., Troncoso, A., Riquelme, J.C. and Ruiz, J.S.A., 2008, December. LBF: A labeled-based forecasting algorithm and its application to electricity price time series. In Data Mining, 2008. ICDM'08. Eighth IEEE International Conference on (pp. 453-461). IEEE.

Martinez Alvarez, F., Troncoso, A., Riquelme, J.C. and Aguilar Ruiz, J.S., 2011. Energy time series forecasting based on pattern sequence similarity. Knowledge and Data Engineering, IEEE Transactions on, 23(8), pp.1230-1243.

